---
title: "Shopping Behavior Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,message=FALSE,warning=FALSE,echo=FALSE}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(recommenderlab)
library(ggplot2)
library(knitr)
library(kableExtra)
library(DT)

# Load datasets
behavior_data <- read.csv("C:/Users/harik/Downloads/shopping_behavior_updated.csv")
trends_data <- read.csv("C:/Users/harik/Downloads/shopping_trends.csv")

# Merge datasets by Customer ID
combined_data <- merge(behavior_data, trends_data, by = "Customer.ID", all = TRUE) %>%
  select(Customer.ID, 
         Gender = Gender.x,  # Include Gender column
         Item.Purchased = Item.Purchased.x, 
         Category = Category.x, 
         Size = Size.x, 
         Color = Color.x, 
         Season = Season.x) %>%
  mutate(Product.Features = paste(Category, Color, Size, Season, sep = " "))

# Create the rating data
rating_data <- combined_data %>%
  group_by(Customer.ID, Item.Purchased) %>%
  summarise(rating = n(), .groups = 'drop')  # Counting the number of purchases

# Relax the filtering criteria if needed (to avoid removing all entries)
rating_data_filtered <- rating_data %>% 
  filter(rating > 0)  # Keep rows with at least 1 purchase

# Create the rating matrix with valid data
rating_matrix <- rating_data_filtered %>%
  pivot_wider(names_from = Item.Purchased, values_from = rating, values_fill = 0)

# Convert to matrix format
rating_matrix <- as.matrix(rating_matrix[, -1])  # Exclude Customer.ID column
rownames(rating_matrix) <- rating_data_filtered$Customer.ID  # Set row names as Customer IDs

# Convert to realRatingMatrix class
rating_matrix <- as(rating_matrix, "realRatingMatrix")

# Define goodRating threshold (e.g., ratings >= 1 are considered good)
good_rating_threshold <- 1  # You can adjust this based on your criteria

# Create an evaluation scheme, specifying goodRating for UBCF
train_test_split <- evaluationScheme(rating_matrix, method = "split", train = 0.8, given = 1, goodRating = good_rating_threshold)

# Run the evaluation
evaluation_results <- evaluate(train_test_split, method = "UBCF", n = c(1, 5, 10))

# Check available metrics in evaluation_results
cat("Available Metrics in Evaluation Results:\n")
print(evaluation_results@results)

# If RMSE or MAE are found, extract and print them
if ("RMSE" %in% colnames(evaluation_results@results)) {
  rmse <- evaluation_results@results$RMSE
  mae <- evaluation_results@results$MAE
  
  # Print RMSE and MAE for each n
  cat("\nRMSE and MAE metrics for different 'n' values:\n")
  rmse_mae_table <- data.frame(n = c(1, 5, 10), RMSE = rmse, MAE = mae)
  print(rmse_mae_table)
  
} else {
  cat("\nNo RMSE or MAE metrics found in the evaluation results.\n")
}

# **1. Popular Items Purchased (Table)**
popular_items <- combined_data %>%
  group_by(Item.Purchased) %>%
  summarise(Purchase.Count = n(), .groups = 'drop') %>%
  arrange(desc(Purchase.Count))

kable(popular_items, caption = "Popular Items Purchased")

# **2. Seasonal Purchases (Table)**
seasonal_data <- combined_data %>%
  group_by(Season) %>%
  summarise(Purchase.Count = n(), .groups = 'drop')

kable(seasonal_data, caption = "Seasonal Purchases") %>%
  kable_styling(full_width = F, position = "center") %>%
  row_spec(0, bold = TRUE, color = "white", background = "#0073e6") %>%
  column_spec(2, color = "red", background = "lightgray") %>%
  add_header_above(c("Purchase Info" = 2))

# **3. Customer Distribution by Gender (Pie Chart)**
gender_data <- combined_data %>%
  group_by(Gender = ifelse(is.na(Gender), "Unknown", Gender)) %>%
  summarise(Purchase.Count = n(), .groups = 'drop') %>%
  mutate(Percentage = round((Purchase.Count / sum(Purchase.Count)) * 100, 1))

# Pie Chart
ggplot(gender_data, aes(x = "", y = Purchase.Count, fill = Gender)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  labs(title = "Customer Distribution by Gender", x = NULL, y = NULL) +
  theme_void() +
  theme(legend.position = "right") +
  scale_fill_brewer(palette = "Set3") +
  geom_text(aes(label = paste0(Gender, ": ", Percentage, "%")), position = position_stack(vjust = 0.5), size = 5)

# **4. Category-wise Purchases (Table & Graph)**
category_purchase_data <- combined_data %>%
  group_by(Category) %>%
  summarise(Purchase.Count = n(), .groups = 'drop')

kable(category_purchase_data, caption = "Category-wise Purchase Count")

ggplot(category_purchase_data, aes(x = reorder(Category, -Purchase.Count), y = Purchase.Count, fill = Category)) +
  geom_bar(stat = "identity") +
  labs(title = "Category-wise Purchase Count", x = "Category", y = "Purchase Count") +
  theme_minimal() +
  coord_flip()

# **5. Visualizing RMSE and MAE for Different n Values (Bar Graph)**
if (exists("rmse_mae_table")) {
  rmse_mae_table %>%
    tidyr::pivot_longer(cols = c("RMSE", "MAE"), names_to = "Metric", values_to = "Value") %>%
    ggplot(aes(x = factor(n), y = Value, fill = Metric)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = "RMSE and MAE for Different n Values", x = "Top-N", y = "Value", fill = "Metric") +
    theme_minimal()
}

# **6. Precision-Recall Tradeoff**
if ("precision" %in% colnames(evaluation_results@results) & "recall" %in% colnames(evaluation_results@results)) {
  precision_recall_data <- data.frame(precision = evaluation_results@results$precision,
                                      recall = evaluation_results@results$recall,
                                      k = c(1, 5, 10))
  
  ggplot(precision_recall_data, aes(x = recall, y = precision, color = factor(k))) +
    geom_point(size = 4) +
    geom_line() +
    labs(title = "Precision-Recall Tradeoff", x = "Recall", y = "Precision", color = "k") +
    theme_minimal()
}

# **7. Visualizing Popular Items Purchased (Bar Chart)**
ggplot(popular_items, aes(x = reorder(Item.Purchased, -Purchase.Count), y = Purchase.Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Popular Items Purchased", x = "Item Purchased", y = "Count") +
  theme_minimal() +
  coord_flip()

# **8. Visualizing Seasonal Purchases (Bar Chart)**
ggplot(seasonal_data, aes(x = Season, y = Purchase.Count, fill = Season)) +
  geom_bar(stat = "identity") +
  labs(title = "Purchases by Season", x = "Season", y = "Count") +
  theme_minimal()

# **Interactive Table with DT**
datatable(popular_items, options = list(pageLength = 5, autoWidth = TRUE))